---
title: "Visualization"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r}
library(tidyverse)
library(ggridges)
```

##Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots!!

Create my first scatterplot ever.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

New approach, same plot.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save and edit a plot object.

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```

## Advanced scatterplot...

Start with the same one and make it fancy!
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() + 
  geom_smooth(se = FALSE)
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) + 
  geom_smooth()
```

Let's facet some things!!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .2) + 
  geom_smooth(se = FALSE, size = 2) +
  facet_grid(name ~ .)
```

Let's combine some elements and try a new plot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5)+
  geom_smooth(se = FALSE)+
  facet_grid(. ~ name)
```

## Some small notes

How many geoms have to exist?

You can have whatever geoms you want.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

You can use a neat geom!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d() +
  geom_point(alpha = .3)
```

## Univariate plots

Histogram are really great.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Can we add color..

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge") +
  facet_grid(. ~ name)
```

Let's try a new geometry

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3)
```

What about box plots??

```{r}
weather_df %>% 
  ggplot(aes(x =name, y = tmin)) +
  geom_boxplot()
```

Trendy plots

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin() +
  stat_summary(fun = "median")
```

Ridge plots -- the most popular plot of 2017

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

## Save and Embed

Let's save a scatterplot

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)

ggsave("./results/weather_plot.pdf", weather_plot, width = 8, height = 5)

```

What about embedding..

```{r}
weather_plot
```

```{r, fig.width=6, fig.height=3}
weather_plot
```

##Visualization II
####Remeber this plot..?
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)

```

## Labels
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)+
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature(C)",
    y = "Minimum daily temperature(c)",
    caption = "Data from rnoda package; temperatures in 2017."
  )
```

## Scales

Start with the same plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)+
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature(C)",
    y = "Minimum daily temperature(c)",
    caption = "Data from rnoda package; temperatures in 2017."
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    position = "right"
  )
```

Look at color scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)+
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature(C)",
    y = "Minimum daily temperature(c)",
    caption = "Data from rnoda package; temperatures in 2017."
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100, 300))
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)+
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature(C)",
    y = "Minimum daily temperature(c)",
    caption = "Data from rnoda package; temperatures in 2017."
  ) +
# Can search for color plate
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)
```

## Themes

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)+
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature(C)",
    y = "Minimum daily temperature(c)",
    caption = "Data from rnoda package; temperatures in 2017."
  ) +
# Can search for color plate
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)+
  theme(legend.position = "bottom")
```

Change the overall theme

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)+
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature(C)",
    y = "Minimum daily temperature(c)",
    caption = "Data from rnoda package; temperatures in 2017."
  ) +
# Can search for color plate
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
# can also use theme_
  ggthemes:: theme_economist()+
  theme(legend.position = "bottom")
```

## Setting options

```{r,eval=FALSE}
library(tidyverse)

knitr:: opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d()
```

